---
pagetitle: "Publications"
output: 
  html_document:
    toc: false
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 4
    css: pubs.css
    includes:
       in_header: [g_analytics.html,header_pubs.html]
---

<style>
    .toc-content {
        padding-left: 10%;
        padding-right: 10%;
    }
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=FALSE)
```

<!--Note: publications are compiled in Rmarkdown directly from a .bib file. Slowly working on making this work. Below is a plot of the semantic similarity between the publications (using the BEAGLE Abstract vectors). Hover over the dots to see the title (still in beta).-->

```{r, message=FALSE,echo=FALSE,warning=FALSE, fig.width=5,fig.height=3, eval=FALSE}
load('semantic_publicatins.RData')
library(plotly)
semantic_plot
```


```{r, results='asis', echo=F, message=F, warning=FALSE, eval=FALSE}
library(bib2df)
bib_df<-bib2df(file="Crump.bib")

library(stringi)

bib_df$TITLE<-stri_replace_all_regex(bib_df$TITLE, "[\\{\\}]", "")
bib_df$JOURNAL<-stri_replace_all_regex(bib_df$JOURNAL, "[\\{\\}]", "")

# sort bib_df by year
bib_df<-bib_df[order(bib_df$YEAR, decreasing=T),]

dims<-dim(bib_df)
start_year <- 0 
for(i in 1:dims[1]){
  a<-bib_df$FILE[i]
  split_a<-unlist(strsplit(a,";"))
  file_path<-paste0("https://github.com/CrumpLab/CrumpLab.github.io/raw/master/",split_a[1])
  if(bib_df[i,]$YEAR != start_year){
    cat(c("##",bib_df[i,]$YEAR, "\n"))
    start_year <- bib_df[i,]$YEAR
  }
  
  cat(c("###",bib_df[i,]$TITLE)) #title
  
  cat(c("\n"))
  
  cat("<div class='row'>")
    cat("<div class='column left'>")
      cat(c("<div id=citationstyle>"))
      cat(c(cat(bib_df[i,]$AUTHOR[[1]],sep=", "), " (",
            bib_df[i,]$YEAR, "). ",
            bib_df[i,]$TITLE, ". *",
            bib_df[i,]$JOURNAL, ", ",
            bib_df[i,]$VOLUME, "*, ",
            bib_df[i,]$PAGES, "."), sep="")
      if(is.na(a)==FALSE){
        cat(c(" ", "[pdf](",file_path,")"), sep="")
      }
      get_data <- bib_df[i,]$NOTE
      if(is.na(get_data)==FALSE){
        cat(c(" | ", get_data), sep="")
      }
      
      cat(c("</div>"))
    
    cat("</div>")
  cat("<div class='column right'>")
    get_doi <- bib_df[i,]$DOI
    if(is.na(get_doi)==FALSE){
      cat("<span class='__dimensions_badge_embed__' data-doi='",get_doi,"' data-style='small_circle' data-legend='hover-left' style = 'text-align:right;'></span>", sep="")
      cat("<span data-badge-popover='left' data-badge-type='donut' data-doi='",get_doi,"' data-hide-no-mentions='true' class='altmetric-embed' style = 'text-align:right;'></span>", sep="")
    }
  
  cat("</div>")
  cat("</div>")
  
  
  
  
  
    
  cat(c("\n"))
  cat(c("\n"))
  cat(c("<div id = abstractstyle>"))
  cat(bib_df[i,]$ABSTRACT)
  cat(c("</div>"))
  cat(c("\n"))
  cat(c("\n"))
  cat(c("\n"))
  
}

# library(dplyr)
# 
# table_df <- bib_df %>%
#             select(YEAR,
#                    TITLE,
#                    JOURNAL,
#                    AUTHOR) %>%
#   mutate(AUTHOR = sapply(AUTHOR, function(x) paste(x,collapse="; ")))
# 
# knitr::kable(table_df)

```


**Copyright Notice:** Some of the documents listed below are available for downloading. These have been provided as a means to ensure timely dissemination of scholarly and technical work on a noncommercial basis. Copyright and all rights therein are maintained by the authors or by other copyright holders, notwithstanding that they have offered their works here electronically. It is understood that all persons copying this information will adhere to the terms and constraints invoked by each author's copyright. These works may not be re-posted without the explicit permission of the copyright holder.

## In progress
 <script src="https://bibbase.org/show?bib=https://crumplab.github.io/Crump_inprogress.bib&jsonp=1&nocache=1&theme=side&authorFirst=1"></script> 

## Published
 <script src="https://bibbase.org/show?bib=https://crumplab.github.io/Crump.bib&jsonp=1&nocache=1&theme=side&authorFirst=1"></script> 

