---
title: "Research opportunities"
output: html_document
---

### Ph.D. students

Students interested in pursuing a Ph.D. with Dr. Crump should apply to the Cognition, Language, and Development training area in the Psychology Doctoral Program at the Graduate Center of CUNY. The deadline for applications December 1st, 2018. [You can apply directly here](http://www.gc.cuny.edu/Prospective-Current-Students/Prospective-Students/Admissions). Learn more about the [Cognition, Language, and Development training area here](http://cunycld.org). Interested applicants should also email Dr. Crump to introduce themselves and their research interests.

### Research experience for Brooklyn College students

We are always looking for students (undergraduate or master's) who are interested in human cognition (how people think, learn, remember, attend, etc.), and who want to gain research experience by learning computational techniques for running experiments and analyzing data.

All students interested in joining the lab should first [complete this entry assignment](#entry), which consists of writing a short computer program. The entry assignment takes a little bit of time and effort. Research takes a lot of time and effort, and completing the assignment will show us that you are motivated to spend time on research. More important, the assignment will help you decide if you will enjoy working on research problems in the lab. If you enjoy the process of completing the assignment, then we think you will look forward to working on research problems in the lab. You should also [read some of the lab publications](https://crumplab.github.io/publications/) to see if you are interested in the research questions we are asking.

### Course credit for Research and Honor's Theses

Most students who join the lab enroll in research classes so that they receive course credit as they gain research experience in the lab. If you are interested in taking one of the courses, please complete the entry assignment above, and then email Dr. Crump about availability. The classes include:

- Psych 2001-4 Laboratory Experience (3 credits each)
- Psych 5001-4 Independent Research (3 credits each)

The Laboratory Experience courses are intended for students who are new to research, and the Independent Research courses are for more senior students. They are each one semester long, and involve conducting research in the lab and writing a research report as a final project.

Several students have completed honor's theses in the lab ([see here for departmental requirements](http://www.brooklyn.cuny.edu/web/academics/schools/naturalsciences/undergraduate/psychology/students/honors.php)). An honor's thesis consists of two semesters of Independent Research (e.g., Psych 5001 and 5002) and culminates in writing a thesis covering the research conducted over the year. The honor's thesis option is a great opportunity for students interested in pursuing research in a Ph.D. program following undergraduate studies. The honor's thesis is a major commitment in time and effort for everyone involved, so students who complete honor's theses in the lab are usually already members of the lab and have previous research experience in the lab (e.g., have taken Psyc 2001 with Dr. Crump). So, if you are interested in research you should plan to get involved as early as possible.

### Undergraduate Research

Recent research projects completed in the lab.

```{r, echo=F}
library(data.table)
undergrad_df<-fread("undergrad_research.csv")
knitr::kable(undergrad_df)

```

### Master's Research

```{r, echo=F}
library(data.table)
masters_df<-fread("masters_research.csv")
knitr::kable(masters_df)

```

---

## Entry Assignment {#entry}

If you’re interested in becoming part of the lab (volunteering, taking a course etc.), the first thing to do is to complete this entry assignment. The assignment consists of writing a small computer program (no previous programming experience necessary, this is an invitation to start learning programming skills).

### Why are we asking you to do this?

A lot of the activities in the lab involve computer programming: during experiments, the timing, stimulus presentation and key press registration are all controlled by programs we write ourselves. When we analyze the data from the experiments, we again do this in programs, written by us, which take the raw data, calculate things such as averages, display results in graphs, and perform statistical tests.

This assignment allows you to demonstrate your interest in this type of activity, and by investing some time in this assignment you can show your commitment to becoming part of the lab.

### Instructions

Your assignment is to write a computer program that does something interesting, then email Dr. Crump with your completed assignment to set up a time to talk about gaining research experience in the lab.

This page gives the instructions for completing the assignment. The files have been produced by various members of the lab, and each of them (with a few exceptions) contains a specific program. We keep all of the programs written by students who entered the lab before you, and your assignment is to write your own program so that we can add your file as well.

Here are your instructions.

1. You will be using the programming language R to complete the assignment (we use this in the lab for data-analysis, and lots of other things).
  * Download and install the most recent version of R for your computer system (it's free). The R website is [https://www.r-project.org](https://www.r-project.org). Once installed you will now have the R programming language on your computer
  * Download and install R-Studio (also free, R must be installed first). The website is [https://www.rstudio.com](https://www.rstudio.com). Once installed you will have a GUI (graphic user-interface) that will help you use R. You will use R-studio to write and run R programs. It provides a text editor for writing programs, a console/terminal for running the programs, and views for looking at variables, and graphs, and many other things.  

2. Familiarize yourself with R-studio by looking at tutorials and example code. There are lots of resources in blogs, and youtube videos, all a google search away.
	* For a general introduction to R and R-studio you can [read chapters 2 and 3 in Dr. Crump's computer programming textbook](https://crumplab.github.io/ProgrammingTextbook/).

3. Run and examine the example programs written by members of the lab.
  * For students with no programming experience, we have created a set of example programs written in R to get you started. You can download and run these scripts in R-studio. The scripts are ordered in terms of difficulty, and contain many comments that explain what each line does.
	* [The example code is stored in this github repository.](https://github.com/CrumpLab/EntryAssignment) You can view the code for each program directly in your web-browser through github. You will need to run these programs in R-studio to see how they work. You can download all of the files to your computer using the "clone or download" button.
  * Examine the code of the programs as you try them. We recommend examining the code of programs 1-5 in that order, because they were designed to go from easier to harder. As you examine these programs, make use of the comments (the text in green, marked by a # at the beginning of each line) that describe what’s going on.

4. Write your own program. Having studied the programs that other people in the lab have written, now it is time to start your assignment: think of a program, any program, that does something interesting and that can be added to the list of programs that members of the lab have created for this assignment. Create whatever program you like, really – as long as it shows that you’ve put in some time to learn a thing or two about R programming.
  * This is what you do: In R-studio, create a new R-Script (file->new file->R-Script, or click the document icon with a green +). Save your script and give it a name. Then, use the text-editor window to code your program. You can try out all of the lines of code in the console, and you can run all of the lines at once by pressing the run button, or by copying and pasting all of the code into console and pressing enter.
	* You should write your program so that it can be loaded into R-studio and run, with the output of program producing some result the accomplishes the goal of the program. For example after running the program, the result might be displayed in the console, or saved as a file, or shown in a graph, you decide (it's your program).
	* If you are looking for suggestions for programs to write you could try one of the programming challenges listed in [chapter three of the programming textbook](https://crumplab.github.io/ProgrammingTextbook/).

5. Email your R-script to Dr. Crump (mcrump@brooklyn.cuny.edu) and request a time to meet about gaining research experience in the lab.
