---
title: "R"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    includes:
       in_header: g_analytics.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE, eval=F)
```

I forget how to do so many things in programming. I mainly use R now, so this is my elephants grave yard of forgotten code that I need to remember for later.

## dplyr

[dplyr cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf)

```{r}
library(dplyr)
```

### select

`select` for picking specific columns

```{r}
all_data   <- mtcars

smaller_df <- all_data %>%
                select(mpg,cyl)


```

### Summarize data

```{r}

summarise(all_data, mean=mean(mpg),
                    sd=sd(mpg))

```

### Group by

```{r}
summary_df <- all_data %>% 
              group_by(cyl,am) %>%
              summarize(mean=mean(mpg),
                        se=sd(mpg)/sqrt(length(mpg)))
```

### Filter rows

```{r}

restricted_df <- all_data %>% 
                 filter(gear > 3)

restricted_df <- all_data %>% 
                 filter(gear > 3,
                        am == 1)

```

### Mutate

Add columns, delete columns

```{r}
dim(all_data)

all_data <- all_data %>% 
            mutate(newCol = rep(0,32))

all_data <- all_data %>% `[<-`("mpg",value=NULL)


```
### Arrange order

```{r}
arrange(mtcars, mpg)  # ascending (small to large)

arrange(mtcars, desc(mpg)) # descending (large to small)

```

## Papaja

For apa style papers

### Figures

1. jou mode doesn't respect fig.width or fig.height, use out.width="3in" for single column
2. use out.width="\\textwidth" and fig.env="figure*" to span both columns

### Indentation in jou mode

for yaml

```{r, eval=F}
header-includes:
   - \setlength\parindent{24pt}
```

### colons in yaml

Use the code `&#58;` for colons

### add latex to header for journal name

add to yaml

```
header-includes:
  - \journal{Journal of Foobar}

```

## kableextra

For the grammar of tables...Makes awesome tables with ease!

[http://haozhu233.github.io/kableExtra/awesome_table_in_html.html#overview](http://haozhu233.github.io/kableExtra/awesome_table_in_html.html#overview)




