---
title: "A reading list on eugenics"
description: "*June 28th, 2022* - This is an annotated bibliography for books I am reading on eugenics and its connections to psychology and society."
author: "Matt Crump"
date: '2022-06-28 `r paste("Last knit: ", lubridate::today())`'
bibliography: EugenicsBooks.bib
opengraph:
  twitter:
    card: summary
    creator: "@MattCrumpLab"
    site: "@MattCrumpLab"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A reading list on eugenics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo=FALSE,
  message= FALSE,
  warning=FALSE
)
```

```{css, results='asis', echo=F   }
@media (min-width: 1400px) {
    body {
        font-size:18px
    }

    .col-md-3 {
        margin-left: 5rem
    }

    .row {
        --bs-gutter-x: -20%!important;
    }

    .row>main {
        max-width: 40rem;
    }

}

.apa {
  margin-left: 2em;
  text-indent: -2em;
}

```

Eugenics and related ideological systems continue to influence psychology and society. I am using this post as an annotated bibliography to gather my reading list about eugenics. The full list of references is at the bottom. I plan to update this list over time as I find more to read on the topic. 

```{r,results="asis"}
crumplab::bib2annotate("yakushkoScientificPollyannaismInquisition2019","EugenicsBooks.bib")
```

One of my favorite books. Critical on many aspects of psychology, a transformational read. Chapters on eugenics are excellent. A large number of early psychologists were also deeply involved in the eugenics movement for well over a half-century.

---

```{r,results="asis"}
crumplab::bib2annotate("kuhlBettermentRaceRise2013","EugenicsBooks.bib")
```

Establishes the international scope and scale, along with shifting ideologies within the eugenics movement.

---

```{r,results="asis"}
crumplab::bib2annotate("bashfordOxfordHandbookHistory2010","EugenicsBooks.bib")
```

Multiple authors contribute chapters. Individual coverage of eugenics in regional/national contexts. Individual coverage of eugenics in societal contexts: e.g., health care, women's rights, science, genetics, etc.

---

```{r,results="asis"}
crumplab::bib2annotate("mclarenOurOwnMaster2014","EugenicsBooks.bib")
```

Excellent history of the eugenics movement in Canada, including thorough linkages between prominent Canadian eugenicists and their influence on major provincial and national institutions.

---

```{r,results="asis"}
crumplab::bib2annotate("prestonDisasterEducationRace2012","EugenicsBooks.bib")
```

Disaster education refers to state produced information related to survival following disasters, such as nuclear Armageddon. This book is really fascinating. The history of disaster education material is deeply mired in eugenics and white supremacy. Chapters also include useful contextual excursions.

---


```{r,results="asis"}
crumplab::bib2annotate("chittyEugenicsRaceIntelligence2007","EugenicsBooks.bib")
```

A historical analysis of the UK education system in terms of eugenics and class struggles.

---

```{r,results="asis"}
crumplab::bib2annotate("currellPopularEugenicsNational2006","EugenicsBooks.bib")
```

Multiple authors contribute chapters on US eugenic culture in the 1930s. Very interesting read, especially the visual eugenics chapters. 

---

```{r,results="asis"}
crumplab::bib2annotate("wilsonEugenicMindProject2017","EugenicsBooks.bib")
```

Standpointing eugenics and focus on the Alberta, Canada context.

---

```{r,results="asis"}
crumplab::bib2annotate("tuckerFundingScientificRacism2007","EugenicsBooks.bib")
```

Title says it all. Fills in many gaps about funding sources for racist science, including much funding for many racist psychologists.

---

```{r,results="asis"}
crumplab::bib2annotate("seldenInheritingShameStory1999","EugenicsBooks.bib")
```

Another cogent introduction to the eugenics movement in the USA, with extra focus on how eugenics set the framework for the American education system.

---



## References









