---
title: "What do we know about the psychology of learning to computer program? I don't know, so I'm doing a lit review..."
author: "Matt Crump"
date: "`r format(Sys.time(), '%d %B, %Y')`"
twitter: "https://twitter.com/MattCrump_"
github: "https://github.com/CrumpLab"
website: "https://crumplab.gihub.io"
bibliography: refs.bib
csl: web/apa.csl
output: 
  html_document:
    template: web/template.html
    toc: true
    toc_float: true
    collapsed: false
    code_folding: show
    number_sections: false
    toc_depth: 4
    theme: yeti
    highlight: kate
    css: web/crump_basic.css
    includes:
      in_header: web/header.html
    md_extensions: -autolink_bare_uris
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

---

How do people learn to program a computer? I've got some opinions, some experience learning myself, and some experience teaching computer programming skills to students in Psychology. But, I'm not up to date on the literature on this topic. Time for a literature review. Let's see how far I can get in a couple hours (or more, let's see).

## Setting up and notes for my future self

I'm hoping that the work I put in now might be useful for my self later on. So, I'm going to blog about doing this literature review while I do it. I've also got some useful tools at my disposal, namely [Zotero](https://www.zotero.org), a free reference manager, that I will use to store the papers the I find, and to create a .bib file that I can use to cite the papers. I'll probably also tweet about this a little bit (already got some nice hits after asking for refs that I should be looking it).

Some time ago I remember wading through some literature on teaching computer programming...I'm sad to see that I didn't save any of that in a Zotero folder...ooops, all that work is now lost to time.

## Some questions that I have right now

1. What kind of assessments are used to measure knowledge of computer programming concepts and skills?

2. Are there good examples of specific exercises or interventions that have a measurable and reliable influence on learning/retaining/understanding a computer programming concept or skill.

3. How is this literature organized? Are researchers primarily interested in the applied problem of "optimizing" curriculum for teaching computer programming, or using the domain of computer programming as a model learning environment to examine aspects of human cognition?

4. Are their tips I can extract from this literature that I should (or should not I suppose) employ in the class room when I teach programming concepts?

5. I'm a cognitive psychologist with a general interest in skill learning...am I interested in wading into this research area as a part of my own research?


## [http://teachtogether.tech](http://teachtogether.tech)

Thanks to [\@mariacdangelo](https://twitter.com/mariacdangelo) for pointing me to [http://teachtogether.tech](http://teachtogether.tech), a CC-licensed book that formed part of the Software Carptentry training program. Lot's to digest here, including a long reference section. I guess I'll start mining this reference section and see where it takes me. 

## Papers

Momentarilly got sucked down a rabbit hole trying to figure out how I want to list things in this .Rmd document. It would be nice to have a function where I could pump in a citation key, and have it print out the full reference so that i can put my notes underneath it. I sort of have this working, but don't like the formatting...Oh, well, I should stop this and continue with the review...

--- 

```{r, results='asis', echo=FALSE}
library(knitcitations)
bib <- read.bibtex("refs.bib")
bib[["bennedsenFailureRatesIntroductory2019"]]
```

@bennedsenFailureRatesIntroductory2019 follow up on their previous paper looking at failure rates in introductory CS1 classes. The previously reported 33% failure rates (in 2007), and are now reporting 28%, a small reduction. One conclusion is the failure rate is not overwhelmingly bad. I remember failing CS1 as an undergrad, so count me in. Too bad CS1 isn't a liberal arts requirement (at least I don't think it is in general). It would be nice to know what the failure rate is for students who don't intend to major in CS.

--- 

```{r, results='asis', echo=FALSE}
bib[["linnCaseCaseStudies1992"]]
```

@linnCaseCaseStudies1992 describe and advocate a "case study" approach. Students are given a problem (e.g., write a program to print block letters given input letters), and are shown "expert" notes on how to organize an algorithm to solve the problem. Methods and results are minimal, but a graph is shown at the end with some data indicating a small benefit to scoring higher on the problem given info about other students solutions along with expert notes. The authors note that students complained about having to read 20 pages worth of expert commentary that was part of the "helpful" learning material, yikes.

I remember some the assignments I was given in an intro CS course, and most of them made no sense to me. In general I've personally found the [project euler](https://projecteuler.net) approach helpful. That website challenges anyone to provide code-based solutions to concrete math problems (e.g., what is the sum of the prime numbers from 1 to 1000), and if you enter the correct answer you get to see other people's answers to the problem in a variety of languages. The idea is that by trying to solve these problems, you end up learning about basic programming constructs like variables, loops, logic, functions etc. All this said, I'm a fan of learning by examples, and maybe I'll keep the "print block letters" problem in my back pocket...I'll also note that I've been creating lists of problems to solve as a means to learning basic programming here [https://crumplab.github.io/programmingforpsych/programming-challenges-i-learning-the-fundamentals.html](https://crumplab.github.io/programmingforpsych/programming-challenges-i-learning-the-fundamentals.html), maybe I should add this one to the list of advanced problems...




## References










